Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-04-19 products/manuscript/Manuscript.qmd
Hopefully works! Pushed manuscript figure captions and worked on tbl err

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "[insert explanation here]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
```{r}
#| message: false
#| label: fig-2
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| message: false
#| label: fig-3
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| message: false
#| label: fig-4
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

>>>>

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "[A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods. The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs.]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
```{r}
#| message: false
#| label: fig-2
#| fig-cap: "[A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies. In this case, a positive correlation is shown. It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| message: false
#| label: fig-3
#| fig-cap: "[Scatter plots are used to see if there is a relationship between two sets of continuous data. In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is being examined.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies. The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa.]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| message: false
#| label: fig-4
#| fig-cap: "[Violin plots are used to visualize the distribution of data across different categories. In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).The violin plot below shows that the distribution of unintended pregnancy rates is wider for countries with high SARC usage compared to counties with medium or low SARC usage. This means that there is more variation in unintended pregnancy rates in counties with high SARC usage.]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```
